---
title: "Sensitivity Power Analysis"
output: html_document
date: "2026-01-06"
---

## Load packages

```{r}
library(readr)
library(readxl)
library(tidyverse)
library(dplyr)
library(sjPlot)
library(easystats)
library(Superpower)
library(lemon)
library(ggeasy)
library(patchwork)
library(svglite)
library(here)

# set plot theme
theme_set(theme_classic(base_size=14, base_family="sans"))
```

## Sensitivity analyses for the 6 outcome measures

Function for calculating pooled standard deviation for multiple samples in one vector.

```{r}
calculate_pooled_sd <- function(sds, ns) {
  # 'sds' is a vector of standard deviations
  # 'ns' is a vector of sample sizes
  
  if (length(sds) != length(ns) || length(sds) < 2) {
    stop("You must provide vectors of SDs and sample sizes for at least two samples.")
  }
  
  # Calculate degrees of freedom for each sample
  dfs <- ns - 1
  
  # Calculate pooled variance
  pooled_variance <- sum((dfs * sds^2)) / sum(dfs)
  
  # Calculate pooled standard deviation
  pooled_sd <- sqrt(pooled_variance)
  
  return(pooled_sd)
}
```

### 10m

First, we need to get the covariate-outcome relationship from Thurlow et al. (2024) (<https://doi.org/10.1007/s40279-023-01959-1>) so we will import the raw data, create a change score and store appropriate study descriptives.

```{r}
(df.10 <- data.frame(
  pre.test = c(1.75, 1.78, 1.86, 1.80, 1.77, 1.88, 1.96, 1.90,
               1.88, 1.87, 1.87, 1.84, 1.73, 1.75, 1.70, 
               1.88, 1.96, 1.90, 1.85, 2.12),
  pre.test.sd = c(0.11, 0.11, 0.13, 0.08, 0.06, 0.1, 0.05,
                  0.07, 0.07, 0.10, 0.09, 0.09, 0.07, 0.05,
                  0.05, 0.21, 0.12, 0.14, 0.14, 0.06),
  post.test = c(1.74, 1.70, 1.82, 1.79, 1.76, 1.86, 1.93, 1.82, 
                1.86, 1.82, 1.85, 1.81, 1.62, 1.64, 1.68,
                1.81, 1.90, 1.82, 1.76, 2.12),
  post.test.sd = c(0.11, 0.12, 0.09, 0.09, 0.06, 0.1, 0.08, 0.06, 
                   0.09, 0.14, 0.11, 0.11, 0.09, 0.07,
                   0.10, 0.17, 0.07, 0.11, 0.11, 0.05),
  study.n = c(18, 18, 10, 10, 13, 8, 7, 12, 21, 9, 13, 14, 8, 7, 9, 
              16, 15, 10, 10, 8)))

# create change scores
(df.10 <- df.10 |> mutate(change = post.test - pre.test))

# 10 m weighted pre-test mean
(m.10 <- df.10$pre.test) # vector for means
(n.10 <- df.10$study.n) # vector for n
(wm.10 <- weighted.mean(m.10, n.10))
(wm.10 <- round(wm.10, 2))

# pooled SD
(sd.10 <- df.10$pre.test.sd) # vector for sds
(n.10 <- df.10$study.n) # vector for n
(pooledsd.10 <- calculate_pooled_sd(sd.10, n.10))
(pooledsd.10 <- round(pooledsd.10, 2))

# covariate-outcome relationship
cor.10 <- correlation(df.10, select = "pre.test", select2 = "change")
(r.10 <- cor.10$r) # save r value
```

#### Sensitivity analysis

```{r}
# set common parameters
n_per_group <- 11
n_groups <- 2 
n_covariates <- 1
sd_val <- pooledsd.10
r2_val <- r.10^2
alpha_level <- 0.05/6

# define a set of 'mu' vectors to test
(mu_scenarios.10 <- list( 
  c(0.00, -0.01), c(0.00, -0.02), c(0.00, -0.03), c(0.00, -0.04),
  c(0.00, -0.05), c(0.00, -0.06), c(0.00, -0.07), c(0.00, -0.08),
  c(0.00, -0.09), c(0.00, -0.10), c(0.00, -0.11), c(0.00, -0.12),
  c(0.00, -0.13), c(0.00, -0.14), c(0.00, -0.15), c(0.00, -0.16),
  c(0.00, -0.17), c(0.00, -0.18), c(0.00, -0.19), c(0.00, -0.20)))

(sen.10m <- tibble(
  Scenario = paste("Scenario", seq_along(mu_scenarios.10)),
  mu_values = mu_scenarios.10) |>
  mutate(Group1.10 = map_dbl(mu_values, ~ .x[1]), # extract individual mu values for clarity
         Group2.10 = map_dbl(mu_values, ~ .x[2])) |>
  mutate(PowerResult = map(mu_values, ~ {  # run the power analysis for each row
    power_oneway_ancova(
      n = rep(n_per_group, n_groups),
      mu = .x,
      n_cov = n_covariates,
      sd = sd_val,
      r2 = r2_val,
      alpha_level = alpha_level,
      type = "exact")})) |>
  mutate(Power = map_dbl(PowerResult, ~ .x$power)) |> # extract only power value
  select(Scenario, Group1.10, Group2.10, Power))   # clean final table

# plot curve 
(ten <- sen.10m |>
  ggplot(aes(Group2.10, Power)) +
  annotate("segment", x = -0.2, xend = -0.18, y = 80, 
             yend = 80, colour = "red", linewidth = 1.5, lineend = "round") + # 80% power
  annotate("segment", x = -0.18, xend = -0.18, y = 0, 
             yend = 80, colour = "red", linewidth = 1.5, lineend = "round") + # critical es
  geom_line(linewidth = 2, lineend = "round") +
  coord_capped_cart(bottom='both', left='both') +
  scale_y_continuous(limits = c(0, 100),
                     breaks = seq(from = 0, to = 100, by = 10)) +
  scale_x_continuous(limits = c(-0.2, 0.0),
                     breaks = seq(from = -0.2, to = 0, by = 0.05)) +
  labs(x = "Change in 10 m Sprint Time (s)", y = "Power (%)")) +
  easy_all_text_colour("black") +
  theme(axis.title.x = element_text(vjust = -1.5)) +
  theme(panel.border=element_blank(), axis.line = element_line()) # to avoid lemon error message
```

### 20m

Repeat previous process with Thurlow et al. (2024) data.

```{r}
(df.20 <- data.frame(
  pre.test = c(2.94, 2.96, 3.20, 3.30, 3.29, 3.20, 3.23,
               3.31, 3.28, 2.96, 3.03, 3.37, 3.28, 3.43),
  pre.test.sd = c(0.11, 0.12, 0.10, 0.09, 0.08, 0.25, 0.14, 
                  0.15, 0.15, 0.1, 0.07, 0.27, 0.25, 0.11),
  post.test = c(2.92, 2.90, 3.20, 3.25, 3.21, 3.22, 3.13,
                3.23, 3.23, 2.85, 2.91, 3.20, 3.14, 3.44),
  post.test.sd = c(0.11, 0.10, 0.10, 0.06, 0.08, 0.22, 0.1,
                   0.21, 0.22, 0.13, 0.11, 0.16, 0.22, 0.08),
  study.n = c(18, 18, 12, 9, 10, 9, 30, 13, 14, 8, 7, 10, 10, 8)))

# Create change scores
(df.20 <- df.20 |> mutate(change = post.test - pre.test))

# get and store Thurlow et al. (2024) study descriptives
# 20 m weighted pre-test mean
(m.20 <- df.20$pre.test) # vector for means
(n.20 <- df.20$study.n) # vector for n
(wm.20 <- weighted.mean(m.20, n.20))
(wm.20 <- round(wm.20, 2))

# pooled SD
(sd.20 <- df.20$pre.test.sd) # vector for sds
(n.20 <- df.20$study.n) # vector for n
(pooledsd.20 <- calculate_pooled_sd(sd.20, n.20))
(pooledsd.20 <- round(pooledsd.20, 2))

# covariate-outcome relationship
cor.20 <- correlation(df.20, select = "pre.test", select2 = "change")
(r.20 <- cor.20$r) # save r value
```

#### Sensitivity analysis

```{r}
# Set common parameters
n_per_group <- 11
n_groups <- 2 
n_covariates <- 1
sd_val <- pooledsd.20
r2_val <- r.20^2
alpha_level <- 0.05/6

# Define a set of 'mu' vectors to test
(mu_scenarios.20 <- list( 
  c(0.00, -0.05), c(0.00, -0.06), c(0.00, -0.07), c(0.00, -0.08),
  c(0.00, -0.09), c(0.00, -0.10), c(0.00, -0.11), c(0.00, -0.12),
  c(0.00, -0.13), c(0.00, -0.14), c(0.00, -0.15), c(0.00, -0.16),
  c(0.00, -0.17), c(0.00, -0.18), c(0.00, -0.19), c(0.00, -0.20),
  c(0.00, -0.21), c(0.00, -0.22), c(0.00, -0.23), c(0.00, -0.24),
  c(0.00, -0.25), c(0.00, -0.26), c(0.00, -0.28), c(0.00, -0.30)))

(sen.20m <- tibble(
  Scenario = paste("Scenario", seq_along(mu_scenarios.20)),
  mu_values = mu_scenarios.20) |>
  mutate(Group1.20 = map_dbl(mu_values, ~ .x[1]), # extract individual mu values for clarity
         Group2.20 = map_dbl(mu_values, ~ .x[2])) |>
  mutate(PowerResult = map(mu_values, ~ {  # run the power analysis for each row
    power_oneway_ancova(
      n = rep(n_per_group, n_groups),
      mu = .x,
      n_cov = n_covariates,
      sd = sd_val,
      r2 = r2_val,
      alpha_level = alpha_level,
      type = "exact")})) |>
  mutate(Power = map_dbl(PowerResult, ~ .x$power)) |> # extract only power value
  select(Scenario, Group1.20, Group2.20, Power))   # clean final table

# plot curve 
(twenty <- sen.20m |>
    ggplot(aes(Group2.20, Power)) +
    annotate("segment", x = -0.3, xend = -0.252, y = 80, 
             yend = 79.9, colour = "red", size = 1.5, lineend = "round") + # 80% power
    annotate("segment", x = -0.252, xend = -0.252, y = 0, 
             yend = 79.9, colour = "red", size = 1.5, lineend = "round") + # critical es
    geom_line(linewidth = 2, lineend = "round") +
    coord_capped_cart(bottom='both', left='both') +
    scale_y_continuous(limits = c(0, 100),
                       breaks = seq(from = 0, to = 100, by = 10)) +
    scale_x_continuous(limits = c(-0.3, 0.0),
                       breaks = seq(from = -0.3, to = 0.0, by = 0.05)) +
    labs(x = "Change in 20 m Sprint Time (s)", y = "Power (%)")) +
  easy_all_text_colour("black") +
  theme(axis.title.x = element_text(vjust = -1.5))
```

### 40m

No 40m data in Thurlow et al. (2024), so here we will use data from Haugen et al. (2020) ([doi: 10.1080/02640414.2020.1741955)](https://www.tandfonline.com/doi/full/10.1080/02640414.2020.1741955).

```{r}
(m1.40 <- 5.45) # forwards
(sd1.40 <- 0.18) # forwards
(n1.40 <- 90) # forwards
(m2.40 <- 5.53) # defenders
(sd2.40 <- 0.16) # defenders
(n2.40 <- 110) # defenders
(m3.40 <- 5.56) # midfielders
(sd3.40 <- 0.17) # midfielders
(n3.40 <- 102) # midfielders

# get and store weighted mean
(m.40 <- c(m1.40, m2.40, m3.40)) # vector for means
(sd.40 <- c(sd1.40, sd2.40, sd3.40)) # vector for sds
(n.40 <- c(n1.40, n2.40, n3.40)) # vector for n
(wm.40 <- weighted.mean(m.40, n.40))
(wm.40 <- round(wm.40, 2))

# pooled SD
(pooledsd.40 <- calculate_pooled_sd(sd.40, n.40))
(pooledsd.40 <- round(pooledsd.40, 2))
```

#### Sensitivity analysis

```{r}
# set common parameters
n_per_group <- 11
n_groups <- 2 
n_covariates <- 1
sd_val <- pooledsd.40
r2_val <- r.20^2
alpha_level <- 0.05/6

# define a set of 'mu' vectors to test
(mu_scenarios.40 <- list( 
  c(0.00, -0.09), c(0.00, -0.10), c(0.00, -0.11), c(0.00, -0.12),
  c(0.00, -0.13), c(0.00, -0.14), c(0.00, -0.15), c(0.00, -0.16),
  c(0.00, -0.17), c(0.00, -0.18), c(0.00, -0.19), c(0.00, -0.20),
  c(0.00, -0.21), c(0.00, -0.22), c(0.00, -0.23), c(0.00, -0.24),
  c(0.00, -0.26), c(0.00, -0.28), c(0.00, -0.29), c(0.00, -0.30),
  c(0.00, -0.31), c(0.00, -0.32), c(0.00, -0.33), c(0.00, -0.34),
  c(0.00, -0.35), c(0.00, -0.36), c(0.00, -0.37), c(0.00, -0.38)))

(sen.40m <- tibble(
  Scenario = paste("Scenario", seq_along(mu_scenarios.40)),
  mu_values = mu_scenarios.40) |>
  mutate(Group1.40 = map_dbl(mu_values, ~ .x[1]), # extract individual mu values for clarity
         Group2.40 = map_dbl(mu_values, ~ .x[2])) |>
  mutate(PowerResult = map(mu_values, ~ {  # run the power analysis for each row
    power_oneway_ancova(
      n = rep(n_per_group, n_groups),
      mu = .x,
      n_cov = n_covariates,
      sd = sd_val,
      r2 = r2_val,
      alpha_level = alpha_level,
      type = "exact")})) |>
  mutate(Power = map_dbl(PowerResult, ~ .x$power)) |> # extract only power value
  select(Scenario, Group1.40, Group2.40, Power))   # clean final table


# plot curve
(forty <- sen.40m |>
    ggplot(aes(Group2.40, Power)) +
    annotate("segment", x = -0.40, xend = -0.285, y = 80, 
             yend = 80, colour = "red", size = 1.5, lineend = "round") + # 80% power
    annotate("segment", x = -0.285, xend = -0.285, y = 0, 
             yend = 80, colour = "red", size = 1.5, lineend = "round") + # critical es
    geom_line(linewidth = 2, lineend = "round") +
    coord_capped_cart(bottom='both', left='both') +
    scale_y_continuous(limits = c(0, 100),
                       breaks = seq(from = 0, to = 100, by = 10)) +
    scale_x_continuous(limits = c(-0.4, -0.05),
                       breaks = seq(from = -0.4, to = -0.05, by = 0.05)) +
    labs(x = "Change in 40 m Sprint Time (s)", y = "Power (%)")) +
  easy_all_text_colour("black") +
  theme(axis.title.x = element_text(vjust = -1.5))
```

## VMax

Calculate the mean and sd of maximal sprint speed in soccer players reported in Table 1 of Haugen et al. (2020) (<https://doi.org/10.1080/02640414.2020.1741955>).

```{r}
# get and store VMax data from Haugen (2020)
(m1.vmax <- 9.3) # forwards
(sd1.vmax <- 0.4) # forwards
(n1.vmax <- 90) # forwards
(m2.vmax <- 9.3) # defenders
(sd2.vmax <- 0.4) # defenders
(n2.vmax <- 110) # defenders
(m3.vmax <- 9.2) # midfielders
(sd3.vmax <- 0.4) # midfielders
(n3.vmax <- 102) # midfielders

# get and store weighted mean
(m.vmax <- c(m1.vmax, m2.vmax, m3.vmax)) # vector for means
(sd.vmax <- c(sd1.vmax, sd2.vmax, sd3.vmax)) # vector for sds
(n.vmax <- c(n1.vmax, n2.vmax, n3.vmax)) # vector for n
(wm.vmax <- weighted.mean(m.vmax, n.vmax))
(wm.vmax <- round(wm.vmax, 2))

# pooled SD
(pooledsd.vmax <- calculate_pooled_sd(sd.vmax, n.vmax))
(pooledsd.vmax <- round(pooledsd.vmax, 2))
```

#### Sensitivity analysis

```{r}
# set common parameters
n_per_group <- 11
n_groups <- 2 
n_covariates <- 1 
sd_val <- pooledsd.vmax
r2_val <- r.20^2 # no covariate-outcome relationship from Thurlow et al., so use 20 m (r.20)
alpha_level <- 0.05/6

# define a set of 'mu' vectors to test
(mu_scenarios.vmax <- list( 
  c(0.00, 0.225), c(0.00, 0.25), c(0.00, 0.275), c(0.00, 0.30),
  c(0.00, 0.325), c(0.00, 0.35), c(0.00, 0.375), c(0.00, 0.40),
  c(0.00, 0.425), c(0.00, 0.45), c(0.00, 0.475), c(0.00, 0.50),
  c(0.00, 0.525), c(0.00, 0.55), c(0.00, 0.575), c(0.00, 0.60),
  c(0.00, 0.625), c(0.00, 0.65), c(0.00, 0.675), c(0.00, 0.70),
  c(0.00, 0.725), c(0.00, 0.75), c(0.00, 0.775), c(0.00, 0.80),
  c(0.00, 0.825), c(0.00, 0.85), c(0.00, 0.875), c(0.00, 0.90)))

(sen.vmax <- tibble(
  Scenario = paste("Scenario", seq_along(mu_scenarios.vmax)),
  mu_values = mu_scenarios.vmax) |>
  mutate(Group1.vmax = map_dbl(mu_values, ~ .x[1]), # extract individual mu values for clarity
         Group2.vmax = map_dbl(mu_values, ~ .x[2])) |>
  mutate(PowerResult = map(mu_values, ~ {  # run the power analysis for each row
    power_oneway_ancova(
      n = rep(n_per_group, n_groups),
      mu = .x,
      n_cov = n_covariates,
      sd = sd_val,
      r2 = r2_val,
      alpha_level = alpha_level,
      type = "exact")})) |>
  mutate(Power = map_dbl(PowerResult, ~ .x$power)) |> # extract only power value
  select(Scenario, Group1.vmax, Group2.vmax, Power))   # clean final table

# plot curve 
(vmax <- sen.vmax |>
    ggplot(aes(Group2.vmax, Power)) +
    annotate("segment", x = 0.2, xend = 0.675, y = 80, 
             yend = 80, colour = "red", size = 1.5, lineend = "round") + # 80% power
    annotate("segment", x = 0.675, xend = 0.675, y = 0, 
             yend = 80, colour = "red", size = 1.5, lineend = "round") + # critical es
    geom_line(linewidth = 2, lineend = "round") +
    coord_capped_cart(bottom='both', left='both') +
    scale_y_continuous(limits = c(0, 100),
                       breaks = seq(from = 0, to = 100, by = 10)) +
    scale_x_continuous(limits = c(0.2, 1.0),
                       breaks = seq(from = 0.2, to = 1.0, by = 0.1)) +
    labs(y = "Power (%)",  x = expression("Change in Vmax (m·s"^{-1}*")")) +
  easy_all_text_colour("black") +
  theme(axis.title.x = element_text(vjust = 1)))
```

## CMJ

To determine other parameters needed for the sensitivity analysis, we can create dataframe with the pre and change values from Thurlow et al. (2024) but using only studies the most common cmj method, i.e., Optojump n = 9 studies.

```{r}
(df.cmj <- data.frame(
  pre.test = c(35.76, 43.87, 47.1, 35.5, 39.96, 29.4, 41.9, 36.6, 34.7),
  pre.test.sd = c(5.26, 6.88, 4.4, 5.8, 5.11, 6.4, 3.8, 4.4, 10.0),
  post.test = c(37.38, 39.0, 49.3, 38.0, 40.86, 30.5, 42.53, 37.33, 35.5),
  post.test.sd = c(5.32, 7.76, 2.6, 7.0, 5.2, 6.8, 3.58, 5.38, 8.6),
  study.n = c(8, 8, 8, 7, 21, 12, 8, 7, 16)))

# create change score
(df.cmj <- df.cmj |> mutate(change = post.test - pre.test))

# get and store weighted pre test mean
(m.cmj <- df.cmj$pre.test) # vector for means
(sd.cmj <- df.cmj$pre.test.sd) # vector for sds
(n.cmj <- df.cmj$study.n) # vector for n
(wm.cmj <- weighted.mean(m.cmj, n.cmj))
(wm.cmj <- round(wm.cmj, 2))

# get pooled pre test sd
(pooledsd.cmj <- calculate_pooled_sd(sd.cmj, n.cmj))
(pooledsd.cmj <- round(pooledsd.cmj, 1))

# covariate-outcome relationship
cor.cmj <- correlation(df.cmj, select = "pre.test", select2 = "change")
(r.cmj <- cor.cmj$r) # save r value
```

#### Sensitivity analysis

```{r}
# set common parameters
n_per_group <- 11
n_groups <- 2 
n_covariates <- 1
sd_val <- pooledsd.cmj
r2_val <- r.cmj^2
alpha_level <- 0.05/6

# define a set of 'mu' vectors to test
(mu_scenarios.cmj <- list( 
  c(0.00, 0.50), c(0.00, 1.00), c(0.00, 1.50), c(0.00, 2.00),
  c(0.00, 2.50), c(0.00, 3.00), c(0.00, 3.50), c(0.00, 4.00),
  c(0.00, 4.50), c(0.00, 5.00), c(0.00, 5.50), c(0.00, 6.00),
  c(0.00, 6.50), c(0.00, 7.00), c(0.00, 7.50), c(0.00, 8.00),
  c(0.00, 8.50), c(0.00, 9.00), c(0.00, 9.50), c(0.00, 10.00), c(0.00, 10.25),
  c(0.00, 10.50), c(0.00, 11.00), c(0.00, 11.50), c(0.00, 12.00),
  c(0.00, 12.50), c(0.00, 13.00), c(0.00, 13.50), c(0.00, 14.00)))

(sen.cmj <- tibble(
  Scenario = paste("Scenario", seq_along(mu_scenarios.cmj)),
  mu_values = mu_scenarios.cmj) |>
  mutate(Group1.cmj = map_dbl(mu_values, ~ .x[1]), # extract individual mu values for clarity
         Group2.cmj = map_dbl(mu_values, ~ .x[2])) |>
  mutate(PowerResult = map(mu_values, ~ {  # run the power analysis for each row
    power_oneway_ancova(
      n = rep(n_per_group, n_groups),
      mu = .x,
      n_cov = n_covariates,
      sd = sd_val,
      r2 = r2_val,
      alpha_level = alpha_level,
      type = "exact")})) |>
  mutate(Power = map_dbl(PowerResult, ~ .x$power)) |> # extract only power value
  select(Scenario, Group1.cmj, Group2.cmj, Power))   # clean final table

# plot curve 
(cmj <- sen.cmj |>
    ggplot(aes(Group2.cmj, Power)) +
    annotate("segment", x = 0, xend = 10.3, y = 80, 
             yend = 80, colour = "red", size = 1.5, lineend = "round") + # 80% power
    annotate("segment", x = 10.3, xend = 10.3, y = 0, 
             yend = 80, colour = "red", size = 1.5, lineend = "round") + # critical es
    geom_line(linewidth = 2, lineend = "round") +
    coord_capped_cart(bottom='both', left='both') +
    scale_y_continuous(limits = c(0, 100),
                       breaks = seq(from = 0, to = 100, by = 10)) +
    scale_x_continuous(limits = c(0.0, 15.0),
                       breaks = seq(from = 0.0, to = 15.0, by = 1.0)) +
  labs(x = "Change in CMJ (cm)", y = "Power (%)")) +
  easy_all_text_colour("black") +
  theme(axis.title.x = element_text(vjust = -1.5))
```

## MAS

To determine parameters needed for the sensitivity analysis, we can create dataframe with descriptive data from Tønnessen et al. (2013) (<https://doi.org/10.1123/ijspp.8.3.323>) and pre-post change scores from Thurlown et al. (2024).

```{r}
(m1.mas <- 16.2) # forwards
(sd1.mas <- 1.0) # forwards
(n1.mas <- 167) # forwards
(m2.mas <- 16.3) # defenders
(sd2.mas <- 0.9) # defenders
(n2.mas <- 237) # defenders
(m3.mas <- 16.4) # midfielders
(sd3.mas <- 0.9) # midfielders
(n3.mas <- 253) # midfielders

# get and store weighted mean
(m.mas <- c(m1.mas, m2.mas, m3.mas)) # vector for means
(sd.mas <- c(sd1.mas, sd2.mas, sd3.mas)) # vector for sds
(n.mas <- c(n1.mas, n2.mas, n3.mas)) # vector for n
(wm.mas <- weighted.mean(m.mas, n.mas))
(wm.mas <- round(wm.mas, 2))
# convert to m.s
(wm.mas <- (wm.mas / 3.6))
(wm.mas <- round(wm.mas, 2))

# pooled SD
(pooledsd.mas <- calculate_pooled_sd(sd.mas, n.mas))
(pooledsd.mas <- round(pooledsd.mas, 2))
# convert to m.s
(pooledsd.mas <- (pooledsd.mas / 3.6)) # convert to m.s
(pooledsd.mas <- round(pooledsd.mas, 2))

# YYIR1 data from Thurlow et al. (2024) 
# Get the covariate-outcome relationship from Thurlow et al. (2024)
(df.yoyo <- data.frame(
  pre.test = c(1105, 1092, 1642, 1686, 2472, 2500, 1917,
               1667, 1792, 2307, 1832, 1029, 1515, 1350,
               1455, 1764, 914, 1830, 1691, 605),
  pre.test.sd = c(314, 238, 365, 359, 223, 246, 440, 441,
                  209, 252, 310, 273, 275, 450, 188, 334,
                  330, 274, 600, 233),
  post.test = c(1435, 1441, 1822, 1811, 2604, 2696, 2455,
                1852, 2065, 2480, 2216, 1303, 1612, 1725,
                1677, 1798, 985, 2270, 2183, 775),
  post.test.sd = c(376, 271, 461, 260, 362, 344, 493, 499,
                   331, 159, 395, 211, 290, 479, 308, 335,
                   337, 294, 645, 242),
  study.n = c(18, 18, 18, 18, 10, 10, 13, 8, 8, 9, 5, 7, 13,
              9, 12, 10, 10, 8, 7, 8)))

(df.yoyo <- df.yoyo |> mutate(change = post.test - pre.test))

# covariate-outcome relationship
plot(cor_test(df.yoyo, "pre.test", "change")) # view relationship
cor.yoyo <- correlation(df.yoyo, select = "pre.test", select2 = "change")
(r.yoyo <- cor.yoyo$r) # save r value
```

#### Sensitivity analysis

```{r}
# set common parameters
n_per_group <- 11
n_groups <- 2 
n_covariates <- 1
sd_val <- pooledsd.mas
r2_val <- r.yoyo^2
alpha_level <- 0.05/6

# define a set of 'mu' vectors to test
(mu_scenarios.mas <- list( 
  c(0.00, 0.10), c(0.00, 0.12), c(0.00, 0.14), c(0.00, 0.16),
  c(0.00, 0.18), c(0.00, 0.20), c(0.00, 0.22), c(0.00, 0.24),
  c(0.00, 0.26), c(0.00, 0.28), c(0.00, 0.30), c(0.00, 0.32),
  c(0.00, 0.34), c(0.00, 0.36), c(0.00, 0.38), c(0.00, 0.40),
  c(0.00, 0.42), c(0.00, 0.44), c(0.00, 0.46), c(0.00, 0.48),
  c(0.00, 0.50), c(0.00, 0.52), c(0.00, 0.54), c(0.00, 0.56),
  c(0.00, 0.58), c(0.00, 0.60), c(0.00, 0.62), c(0.00, 0.64)))

(sen.mas <- tibble(
  Scenario = paste("Scenario", seq_along(mu_scenarios.mas)),
  mu_values = mu_scenarios.mas) |>
  mutate(Group1.mas = map_dbl(mu_values, ~ .x[1]), # extract individual mu values for clarity
         Group2.mas = map_dbl(mu_values, ~ .x[2])) |>
  mutate(PowerResult = map(mu_values, ~ {  # run the power analysis for each row
    power_oneway_ancova(
      n = rep(n_per_group, n_groups),
      mu = .x,
      n_cov = n_covariates,
      sd = sd_val,
      r2 = r2_val,
      alpha_level = alpha_level,
      type = "exact")})) |>
  mutate(Power = map_dbl(PowerResult, ~ .x$power)) |> # extract only power value
  select(Scenario, Group1.mas, Group2.mas, Power))   # clean final table

# plot curve 
(mas <- sen.mas |>
    ggplot(aes(Group2.mas, Power)) +
    annotate("segment", x = 0.05, xend = 0.44, y = 80, 
             yend = 80, colour = "red", size = 1.5, lineend = "round") + # 80% power
    annotate("segment", x = 0.44, xend = 0.44, y = 0, 
             yend = 80, colour = "red", size = 1.5, lineend = "round") + # critical es
    geom_line(linewidth = 2, lineend = "round") +
    coord_capped_cart(bottom='both', left='both') +
    scale_y_continuous(limits = c(0, 100),
                       breaks = seq(from = 0, to = 100, by = 10)) +
    scale_x_continuous(limits = c(0.05, 0.70),
                       breaks = seq(from = 0.05, to = 0.70, by = 0.1)) +
    labs(y = "Power (%)",  x = expression("Change in MAS (m·s"^{-1}*")")) +
  easy_all_text_colour("black") +
  theme(axis.title.x = element_text(vjust = -0.5)))
```

## Save intermediate results for plotting in figures.Rmd

```{r}
# Create list with all sensitivity analysis data
sensitivity_data <- list(
  # Sensitivity analysis dataframes
  sen.10m = sen.10m,
  sen.20m = sen.20m,
  sen.40m = sen.40m,
  sen.vmax = sen.vmax,
  sen.cmj = sen.cmj,
  sen.mas = sen.mas,
  # Critical effect sizes (80% power threshold) - extracted from plot annotations
  critical.10 = -0.18,
  critical.20 = -0.252,
  critical.40 = -0.285,
  critical.vmax = 0.675,
  critical.cmj = 10.3,
  critical.mas = 0.44
)

# Save to .rds file
saveRDS(sensitivity_data, here("data", "sensitivity_analysis.rds"))
```

## Combine plots and save

```{r}
(ten | twenty) / (forty | cmj) / (vmax | mas) & theme(text = element_text(size = 14))

# Save figure
ggsave(here("figures", "figure 2.svg"), dpi = 300, width = 25, height = 25, units = "cm")
```
